Trajectories

### 2023-04-28 docs/index.qmd
Add appendix for syllable frequencies and new revisions

```{r}
#| label: fig-participants-age
#| fig-width: 7
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = clrs[1]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r}
#| label: fig-participants-lp
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
participants %>% 
    pivot_longer(
        c(doe_catalan, doe_spanish),
        names_to = "language", values_to = "doe", 
        names_transform = ~str_to_sentence(str_remove(., "doe_"))
    ) %>% 
    mutate(
        doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
        id = paste0(id, " (", time, ")")
    ) %>% 
    count(language, doe) %>% 
    ggplot(aes(doe, y = n, label = n, fill = language)) +
    geom_col(colour = "white", size = 1) +
    geom_text(position = position_nudge(y = 5)) +
    facet_wrap(~language, ncol = 1) + 
    labs(
        x = "Degree of exposure to Catalan",
        y = "Number of participants",
        colour = "Language",
        fill = "Language"
    ) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(
        legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank()
    )
```

@@

<!-- # ```{r} -->
<!-- #| label: tbl-participants-edu -->
<!-- #|  tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)." -->
<!-- # participants %>% -->
<!-- #     left_join(bvq_data$logs) %>% -->
<!-- #     mutate( -->
<!-- #         edu_parent = fct_explicit_na(as.factor(edu_parent)), -->
<!-- #         doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE) -->
<!-- #     ) %>% -->
<!-- #     count(doe_catalan, edu_parent) %>% -->
<!-- #     right_join( -->
<!-- #         expand_grid( -->
<!-- #             doe_catalan = levels(.$doe_catalan), -->
<!-- #             edu_parent = unique(bvq_data$logs$edu_parent) -->
<!-- #         ) -->
<!-- #     ) %>% -->
<!-- #     rename(total = n) %>% -->
<!-- #     pivot_wider(names_from = edu_parent, -->
<!-- #                 values_from = total,  -->
<!-- #                 values_fill = 0) %>% -->
<!-- #     mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% -->
<!-- #     rename(Total = total) %>% -->
<!-- #     select(-"NA") %>%  -->
<!-- #     gt()  -->
<!-- # summary_rows( -->
<!-- #     columns = Complementary:Primary, -->
<!-- #     fns = list(Total = ~sum(., na.rm = TRUE)), -->
<!-- #     formatter = fmt_integer -->
<!-- # ) %>% -->
<!-- # cols_label(doe_catalan = "DoE Catalan") %>% -->
<!-- # sub_missing(columns = 2:8, missing_text = "--") %>% -->
<!-- # tab_spanner("Educational attainment", Secondary:Primary) %>% -->
<!-- # tab_style( -->
<!-- #     style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")), -->
<!-- #     locations = cells_body(columns = 9, rows = 1:5) -->
<!-- # ) %>% -->
<!-- # tab_style( -->
<!-- #     cell_text(style = "italic"), -->
<!-- #     cells_column_labels(columns = 2:7) -->
<!-- # ) %>% -->
<!-- # tab_style( -->
<!-- #     cell_text(weight = "bold"), -->
<!-- #     cells_column_labels(columns = c(1, 8)) -->
<!-- # ) %>% -->
<!-- # tab_style( -->
<!-- #     cell_text(weight = "bold"), -->
<!-- #     cells_column_spanners(spanners = "Educational attainment") -->
<!-- # ) -->
<!-- ``` -->

@@

```{r}
#| label: tbl-items-summary
#| tbl-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
# items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))
# 
# items %>%
#     unnest(list) %>%
#     group_by(language, list) %>%
#     summarise_at(
#         vars(freq, n_phon, lv),
#         list(
#             n = ~sum(!is.na(.)),
#             mean = ~mean(., na.rm = TRUE),
#             sd = ~sd(., na.rm = TRUE),
#             min = ~min(., na.rm = TRUE),
#             max = ~max(., na.rm = TRUE)
#         )
#     ) %>%
#     select(-c(n_phon_n, lv_n)) %>%
#     rename(n = freq_n) %>%
#     gt(groupname_col = "language", rowname_col = "list") %>%
#     fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>%
#     fmt_percent(matches("lv_"), decimals = 1) %>%
#     tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>%
#     tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>%
#     tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>%
#     cols_merge_uncert(col_val = freq_mean, col_uncert = freq_sd) %>%
#     cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>%
#     cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>%
#     cols_merge_range(col_begin = freq_min, col_end = freq_max) %>%
#     cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>%
#     cols_merge_range(col_begin = lv_min, col_end = lv_max) %>%
#     cols_label(
#         list = "",
#         n = md("*N*"),
#         freq_mean = md("Mean ± SD"),
#         freq_min = md("Range"),
#         n_phon_mean = md("Mean ± SD"),
#         n_phon_min = md("Range"),
#         lv_mean = md("Mean ± SD"),
#         lv_min = md("Range")
#     ) %>%
#     tab_style(
#         cell_text(style = "italic"),
#         cells_column_labels(everything())
#     ) %>%
#     tab_style(
#         cell_text(weight = "bold"),
#         cells_column_spanners(everything())
#     ) %>%
#     summary_rows(
#         fns = list("Grand mean" = ~mean(.)),
#         columns = ends_with("_mean")
#     ) %>%
#     summary_rows(
#         fns = list("N" = ~sum(.)),
#         columns = n,
#         formatter = fmt_integer
#     )
```

@@

```{r}
#| label: fig-items-freq
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 4
#| fig-cap: "Distribution of lexical frequency across items. Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations. Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency. The black dot and interval indicate the mean and standard deviation."
mean_sd <- function(x, ...) {
    m <- mean(x, ...)
    s <- sd(x, ...)
    data.frame(mean = m, ymin = m-s, ymax = m+s)
}

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    ggplot(aes(x = freq)) +
    geom_histogram(fill = clrs[4], colour = "white", bins = 30)  +
    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        legend.title = element_blank()
    )
```

### \# Phonemes

Word length ($Phonemes$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.

```{r}
#| label: fig-items-nphon
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of number of phonemes across items"
items %>%
    mutate(language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish")) %>%
    count(language, n_phon) %>%
    ggplot() +
    aes(
        x = n_phon,
        y = n,
        fill = language,
        colour = language,
        shape = language
    ) +
    facet_wrap(~language, ncol = 1) +
    geom_col(colour = "white") +
    geom_text(aes(label = n), position = position_nudge(y = 5), colour = "black") +
    labs(
        x = "# Phonemes",
        y = "Number of words",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_x_continuous(breaks = 1:15) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(
        legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(),
        legend.title = element_blank()
    )
```

### Phonological similarity (Levenshtein)

@@

```{r}
#| label: fig-items-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = clrs[4], colour = "white") +
    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
              position = position_nudge(y = 3), size = 3.5) +
    labs(x = "Levenshtein distance", y = "Number of words") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

### Frequency by \# Phonemes

```{r}
#| label: fig-items-frequency-phonemes
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq) %>% 
    distinct(te, .keep_all = TRUE) %>% 
    count(n_phon, freq) %>% 
    ggplot(aes(n_phon, freq)) +
    geom_point(size = 2, alpha = 0.5, colour = clrs[4], shape = 1, stroke = 1,
               position = position_jitter(width = 0.1), seed = 888) +
    geom_smooth(method = "lm", formula = "y ~ x" ,
                colour = clrs[1], fill = clrs[1]) +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv, size = n)) +
    geom_point(size = 2, shape = 1, stroke = 1, colour = clrs[4]) +
    geom_smooth(method = "lm", formula = "y ~ x", colour = clrs[1], fill = clrs[1]) +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

### \# Phonemes by Levenshtein

```{r}
#| label: fig-items-nphon-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1, stroke = 1, colour = clrs[4]) +
    geom_smooth(method = "lm", formula = "y ~ x", size = 1,
                colour = clrs[1], fill = clrs[1]) +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-exposure-distribution
#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
#| warnings: false
#| fig-height: 7
#| fig-width: 13
responses %>%
    mutate(
        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
        exposure = cut(
            exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE
        )
    ) %>%
    count(exposure, language) %>%
    ggplot(aes(exposure, n, fill = language)) +
    facet_wrap( ~ language) +
    geom_col(colour = "white") +
    labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
         y = "Number of observations",
         fill = "Language") +

    responses %>%
    mutate(
        language = ifelse(grepl("cat_", item), "Catalan", "Spanish"),
        exposure = cut(
            exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE
        ),
        freq = cut(freq, breaks = seq(0, 9, 0.25), include.lowest = TRUE)
    ) %>%
    count(exposure, freq, language) %>%
    ggplot(aes(
        x = freq,
        y = exposure,
        colour = language,
        size = n
    )) +
    facet_wrap( ~ language) +
    geom_point() +
    labs(
        x = "Frequency (Zipf score)",
        y = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
        colour = "Language",
        fill = "Language",
        size = "Number of observations"
    ) +
    theme(panel.grid = element_blank()) +
    plot_layout(ncol = 1) &
    scale_colour_manual(values = clrs[c(1, 4)]) &
    scale_fill_manual(values = clrs[c(1, 4)]) &
    guides(fill = "none", colour = "none") &
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank())
```

>>>>

```{r}
#| label: fig-participants-age
#| fig-width: 7
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = "grey60") +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(x = "Age (months)",
         y = "# participants") +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55),
                       breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r apatb-participants}
#| label: tab-participants
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about language degree of exposure (DoE) was provided by families before filling the questionnaire. In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish. A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish. A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants |>
    mutate_at(vars(starts_with("doe_")), 
              (\(x) floor(x * 10) / 10)) |> 
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%",
                       "25-50%", 
                       "50-75%",
                       "75-100%"),
            include.lowest = TRUE
        ) |>
            as.character()
    ) |>
    count(age, doe_catalan) |>
    arrange(desc(doe_catalan)) |>
    pivot_wider(names_from = age, values_from = n) |>
    gt() |>
    summary_rows(
        columns = 2:7,
        fns = list("Total" = ~ sum(., na.rm = TRUE)),
        formatter = fmt_integer
    ) |>
    sub_missing(everything(), everything(), missing_text = "--") |>
    cols_label(doe_catalan = "Catalan exposure") |>
    tab_spanner("Age (months)", 2:8) |>
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

@@

```{r}
#| label: tbl-participants-edu
#|  tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)."
participants %>%
    mutate(id = as.character(id)) |> 
    left_join(bvq_data$logs) %>%
    mutate(
        edu_parent = edu_parent |> 
            as.factor() |> 
            fct_na_value_to_level(),
        doe_catalan = cut(doe_catalan,
                          breaks = seq(0, 1, 0.2),
                          include.lowest = TRUE)
    ) %>%
    count(doe_catalan, edu_parent) %>%
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(bvq_data$logs$edu_parent)
        )
    ) %>%
    rename(total = n) %>%
    pivot_wider(names_from = edu_parent,
                values_from = total,
                values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>%
    rename(Total = total) %>%
    select(-"NA") %>%
    gt() |> 
    summary_rows(columns = Complementary:Primary,
                 fns = list(Total = ~sum(., na.rm = TRUE)),
                 formatter = fmt_integer) %>%
    cols_label(doe_catalan = "DoE Catalan") %>%
    sub_missing(columns = everything(),
                missing_text = "--") %>%
    tab_spanner("Educational attainment", 
                Secondary:Primary) %>%
    tab_style(
        style = list(cell_borders(weight = px(2),
                                  sides = "left", 
                                  color = "grey50")),
        locations = cells_body(columns = 9,
                               rows = 1:5)
    ) %>%
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

@@

```{r}
#| label: tbl-items-summary
#| tbl-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(bvq_data$pool, te, item, language, class, semantic_category))

items %>%
    unnest(list) %>%
    summarise(n = n(),
              across(c(freq, n_phon, lv),
                     lst(mean, sd, min, max)),
              .by = c(language, list)) |> 
    gt(groupname_col = "language", 
       rowname_col = "list") %>%
    fmt_number(-c(list, n),
               decimals = 1) %>%
    tab_spanner("Frequency (Zipf)",
                columns = starts_with("freq_")) %>%
    tab_spanner("Length (Phonemes)",
                columns = starts_with("n_phon_")) %>%
    tab_spanner("Levenshtein (cognateness)", 
                columns = starts_with("lv_")) %>%
    cols_merge_uncert(col_val = freq_mean,
                      col_uncert = freq_sd) %>%
    cols_merge_uncert(col_val = n_phon_mean,
                      col_uncert = n_phon_sd) %>%
    cols_merge_uncert(col_val = lv_mean, 
                      col_uncert = lv_sd) %>%
    cols_merge_range(col_begin = freq_min, 
                     col_end = freq_max) %>%
    cols_merge_range(col_begin = n_phon_min, 
                     col_end = n_phon_max) %>%
    cols_merge_range(col_begin = lv_min,
                     col_end = lv_max) %>%
    cols_label(list = "",
               n = md("*N*"),
               freq_mean = md("Mean ± SD"),
               freq_min = md("Range"),
               n_phon_mean = md("Mean ± SD"),
               n_phon_min = md("Range"),
               lv_mean = md("Mean ± SD"),
               lv_min = md("Range")) %>%
    tab_style(cell_text(style = "italic"),
              cells_column_labels(everything())) %>%
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners(everything())) %>%
    summary_rows(fns = list("Grand mean" = ~mean(.)),
                 columns = ends_with("_mean")) %>%
    summary_rows(fns = list("N" = ~sum(.)),
                 columns = n,
                 formatter = fmt_integer)
```

@@

```{r}
#| label: fig-items-freq
#| echo: false
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Distribution of lexical frequency across items. Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations. Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency. The black dot and interval indicate the mean and standard deviation."
mean_sd <- function(x, ...) {
    m <- mean(x, ...)
    s <- sd(x, ...)
    data.frame(mean = m, ymin = m-s, ymax = m+s)
}

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    ggplot(aes(x = freq)) +
    geom_histogram(fill = "grey60",
                   colour = "white",
                   bins = 30)  +
    labs(x = "Lexical frequency (Zipf score)", y = "Number of words") +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        legend.title = element_blank()
    )
```

### Word length

Word length ($Length$), computed as the number of phonemes in the phonological transcription of the word-form in [SAMPA](https://en.wikipedia.org/wiki/SAMPA) format.

```{r}
#| label: fig-items-nphon
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of number of phonemes across items"
items %>%
    count(language, n_phon) %>%
    ggplot(aes(x = n_phon, y = n)) +
    facet_wrap(~language, ncol = 1) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = n),
              position = position_nudge(y = 5),
              colour = "black") +
    labs(x = "Phonemes",
         y = "Number of words",
         fill = "Language",
         shape = "Language",
         colour = "Language") +
    scale_x_continuous(breaks = 1:15) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks.x = element_line(),
          legend.title = element_blank())
```

### Levenshtein (cognateness)

@@

```{r}
#| label: fig-items-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n),
           labels = paste0(
               n, " (",
               percent(prop, accuracy = 0.1), 
               ")")) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = labels),
              position = position_nudge(y = 3), 
              size = 3.5) +
    labs(x = "Levenshtein distance", y = "Number of words") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

### Frequency by word length

```{r}
#| label: fig-items-frequency-phonemes
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq) %>% 
    distinct(te, .keep_all = TRUE) %>% 
    count(n_phon, freq) %>% 
    ggplot(aes(n_phon, freq)) +
    geom_point(size = 2,
               alpha = 0.5,
               colour = "grey60", 
               shape = 1, 
               stroke = 1,
               position = position_jitter(width = 0.1)) +
    geom_smooth(method = "lm",
                formula = "y ~ x" ,
                colour = "black") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv)) +
    geom_point(size = 2,
               shape = 1, 
               colour = "grey60",
               stroke = 1) +
    geom_smooth(method = "lm", 
                formula = "y ~ x", 
                colour ="black") +
    labs(y = "Levenshtein similarity", 
         x = "Lexical frequency (Zipf score)", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

### Word length by Levenshtein

```{r}
#| label: fig-items-nphon-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1,
               stroke = 1,
               colour = "grey60") +
    geom_smooth(method = "lm",
                formula = "y ~ x",
                size = 1,
                colour = "black") +
    labs(x = "# Phonemes", 
         y = "Levenshtein similarity", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank()) 
```

@@

```{r}
#| label: fig-exposure-distribution
#| fig-cap: "Relationship between lexical frequency and language exposure-weighted lexical frequency"
#| warnings: false
#| fig-height: 7
#| fig-width: 13
responses %>%
    mutate(
        exposure = cut(exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE)) %>%
    count(exposure, language) %>%
    ggplot(aes(exposure, n)) +
    facet_wrap( ~ language) +
    geom_col(colour = "white",
             fill = "grey60") +
    labs(x = "Exposure-weighted frequency\n(Frequency \u00d7 DoE)",
         y = "Number of observations",
         fill = "Language") +

    responses %>%
    mutate(exposure = cut(exposure,
            breaks = seq(0, 9, 1),
            include.lowest = TRUE),
        freq = cut(freq, 
                   breaks = seq(0, 9, 0.25), 
                   include.lowest = TRUE)) %>%
    count(exposure, freq, language) %>%
    ggplot(aes(x = freq,
               y = exposure,
               size = n)) +
    facet_wrap( ~ language) +
    geom_point(colour = "grey60",
               alpha = 1/2) +
    labs(x = "Frequency (Zipf score)",
         y = "Exposure-weighted frequency\\
         (Frequency \u00d7 DoE)",
         colour = "Language",
         fill = "Language",
         size = "Number of observations") +
    theme(panel.grid = element_blank()) +
    plot_layout(ncol = 1) &
    scale_colour_manual(values = clrs[c(1, 4)]) &
    scale_fill_manual(values = clrs[c(1, 4)]) &
    guides(fill = "none", colour = "none") &
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.position = "none")
```
